Co-evolution of lexical meaning \& pragmatic use

### 2018-01-12 paper/cogsci-submission2.tex
revision draft 2/3

\begin{example}
  Consider a simple and abstract coordination game. Agents are of two types: positive or negative. If
  agents of different type interact with each other, they obtain a payoff of $0$. If negative meets
  negative, each receives a payoff of $1$. If positive meets positive, they get a payoff of $2$.  A population state is completely characterized by the proportion $x$ of negatives. The fitness
  of negatives in population state $x$ is $f_n(x) = x$, that of positives is $f_p(x) = 2-2x$.
  The average fitness is $\Phi(x) = x f_n(x) + (1-x) f_p(x) = 3x^2 - 4x + 2$. Without mutation,
  the replicator dynamic will update $x$ to $\text{RD}(x) = \nicefrac{f(x)x}{\Phi(x)} = \nicefrac{x^2}{\Phi(x)}$. The update
  function $\text{RD}(x)$ of the replicator step is plotted in Fig.~\ref{fig:Updates_RMS} as the
  blue line. Rest points, for which $\text{RD}(x)=x$, are at $x=0$, $x=1$ and $x= \nicefrac{2}{3}$.
  The former are attractors as nearby points converge to them. Points near $x=\nicefrac{2}{3}$
  also move towards 0 or 1. This is schematically pictured in the topmost phase portrait in
  Fig.~\ref{fig:Phase_RD}. Adding mutation changes the dynamic and its rest points. Let us assume that $Q_{ji} = .9$
  when $j=i$. This is the proportion of types that are replicated faithfully. Conversely, a proportion of $.1$ will change their type from positive to negative, and vice-versa. The update effect of mutation on its own are described by $\text{M}(x) = .9 x + .1 (1-x) = .8x + .1$, plotted as the linear green line in Fig.~\ref{fig:Updates_RMS}. As shown in Figure \ref{fig:Phase_RD}, in this example mutation alone has only one stable rest point. It is located at $x = .5$. If we first take the replicator step and
  then the mutation step in sequence, we obtain the replicator mutator dynamic
  $\text{RMD}(x) = \text{M}(\text{RD}(x)) = \nicefrac{.9x^2 - .2x +.2}{3x^2 - 4x + 2}$, which is plotted in red in
  Fig.~\ref{fig:Updates_RMS}. The rest points are at $x=.121$, $x=.903$ and $x=.609$. The
  former two are attractors (see Fig.~\ref{fig:Phase_RD}).
\end{example}

@@

\subsection{Example}
Section~\ref{sec:si-case-study} shows how these components and their interaction can shed light on the evolution of scalar implicatures. Before moving on, we will illustrate the model's workings in a constrained type space in order to showcase the effect that pressure for communicative success and learnability have on types. To this end, we focus only on the four types from Section~\ref{sec:languages+use}. These types are made up of a combination of either lexicon $\Llack$ or $\Lbound$ and either literal (level-$0$) or pragmatic (level-$1$) behavior.

@@

A population will be made up of frequencies of these four types. We may think about such a population as a point in a two-dimensional space. In Figure \ref{fig:sketch} this space is outlined by the grey dotted square. The first dimension determines the proportion of types that lexicalize ambiguous $\msome$ (the $x$-axis in Figure \ref{fig:sketch}). That is, the proportion of types with $\Llack$. The second dimension determines the proportion of pragmatic language users (the $y$-axis in Figure \ref{fig:sketch}). Populations at the either corner of this space are monomorphic, consisting purely of a single type. If we move from away from edges toward the center of the space, then this two-dimensional representation underspecifies the proportion of types found in a population. The space's center, for instance, could be composed 50\% of pragmatic $\Llack$ and of an equal proportion of literal $\Lbound$; but it could equally well be a population that has each type with the same frequency. This will be sufficient for our purposes, as we are only interested in showcasing the rough trajectories of populations through this space after replicator or mutator steps. 
%The space's corners (and edges) are unequivocal: the lower-left corner corresponds to an population of only literal $\Lbound$ users; the lower-right corner to all pragmatic $\Lbound$ users; the upper-left to literal $\Llack$ users only; and the upper-right corner to the majority all pragmatic $\Llack$ users.

From Section~\ref{sec:languages+use}, we already know some things about this type space. When it comes to communicative success, types with $\Lbound$ have a functional advantage over $\Llack$; literal $\Lbound$ has a slight advantage over pragmatic $\Lbound$; all have a sizable advantage over literal $\Llack$. The red arrows in Figure \ref{fig:sketch} sketch out these relationships. They depict the rough trajectories we may expect if there is only pressure for communicative success. How much each replicator step moves a population ultimately depends on the rationality parameter $\lambda$. In particular, functional differences decrease as $\lambda$ increases: if $\lambda$ is sufficiently high, all types except for literal $\Llack$ speakers will show a strong tendency to associate $\ssome$ with $\msome$ and $\sall$ with $\mall$. As a consequence, when it comes to learnability and transmission fidelity, the likelihood that any of the three remaining types produces a given datum $d$ will be approximately equal. In other words, the likelihood offers little differentiation for learners if $\lambda$ is high. We can therefore expect that, if the prior is flat, no type is passed on very faithfully. This would lead populations to gravitate toward the space's center after mutator steps. This is sketched by the blue dash-dotted arrows in Figure~\ref{fig:sketch}. Neither prior nor posterior alone, but their combination, are what determines the effects of (iterated) Bayesian learning. If, instead of a flat prior, there is a prior that favors, say, $\Llack$ over $\Lbound$ we can expect mutator steps to be skewed toward the upper-half of the space. This is suggested by the purple arrows in Figure~\ref{fig:sketch}. Taking stock, the effects of iterated learning will depend not only on $\lambda$, affecting the likelihood, but also on the prior, how strong the bias is, and the degree to which learners tend to maximize the posterior, controlled by parameter $l$.

Figure \ref{fig:quiver} shows actual trajectories, showcasing how they are modulated by parameter values. The replicator steps make the functional disadvantage of $\Llack$ against $\Lbound$ apparent, particularly for literal types and/or low $\lambda$. They also highlight that the functional difference between literal and pragmatic $\Lbound$ is small. Under functional pressure only, populations move toward the lower part of the square but the steps become small as the edge, representing populations of only $\Llack$, are reached. As for the mutator diagrams, there are two main things to note. As expected, a small bias for a type will lead to larger proportions to adopt it provided that likelihood offers little to no differentiation between types. This is the situation the right-most space in Figure \ref{fig:quiver} depicts. However, the bias is far from absolute. If, instead, certain types are more likely to be recovered based on the linguistic input they produce, they will be passed on more faithfully. This is why, for instance, populations move toward a majority of $\Llack$-types for $\lambda = 5$ and $l = 10$.

@@

Figures \ref{fig:sketch} and \ref{fig:quiver} depict replicator and mutator steps in isolation to better showcase their effects. Their combination will expectedly yield different outcomes, which we now turn to in a case study on scalar implicatures.

@@

The model of the previous section formalizes the evolutionary competition between different
sets of lexical representations and ways of using them. This section looks at a case study on
scalar implicatures. It engages in a formal thought experiment to address the
question: if a population of language users could freely combine different lexica with
different pragmatic strategies, what are conditions under which the majority view of scalar
implicatures could have evolved?

@@

Consider a state space with three states $\States = \set{\snone, \ssome, \sall}$ and think of
it as a partition of possible worlds into cells where none, some or all of the $A$s are $B$s,
for some arbitrary fixed predicates $A$ and $B$. Eight lexical representations can be distinguished based on
their truth or falsity in three world states, six of which are not contradictory or
tautological (see Table~\ref{tab:concepts} below).

@@

Recall that types are a combination of a lexicon and a manner of language use. We analyze the
model's predictions in populations of types with one of the two behaviors introduced earlier;
literal or pragmatic. The former correspond to level $0$ reasoners and the latter to level
$1$. Accordingly, we consider a total of $432$ types. Six are variants of pragmatic language
users with $\Llack$-like lexica. We refer to these as \emph{target types}, because they
represent lexica and language use that conform to the majority view of scalar
implicatures. Twelve types are either literal or pragmatic speakers with lexica of the
$\Lbound$ kind. We refer to these as \emph{competitor types}, because they are expected to be
the target types' main contenders in evolutionary competition. Finally, note that while
different types may lexicalize the same representations, they may nevertheless map different
states to different overt messages. More informally, they speak different languages that
lexicalize the same concepts. Consequently, more often than not different types of the same
kind fail to understand each other.

@@

\hl{The discussion on derivation costs is not very satisfying on my opinion. For instance, the condition of "simple representations being favored over more complex ones" is compatible with many more cost specifications.}

There is a growing effort to develop empirically testable representational languages that allow
for the measure of semantic complexity. For instance, so-called {\em languages of thought}
(LOTs) have been put to test in various rational probabilistic models that show encouraging
results (see, e.g., \citealt{katz+etal:2008}, \citeauthor{piantadosi+etal:underreview} under review; 2012, and \citealt{piantadosi+jacobs:2016} for recent discussion). At its
core, a LOT defines a set of operations and composition rules from which lexical representations can be
derived. As a first approximation and for the sake of concreteness, we follow this approach to
motivate and formalize a preference of learners for simpler semantic representations
(\citealt{feldman:2000, chater+vitanyi:2003, piantadosi+etal:2012a,kirby+etal:2015}, \citeauthor{piantadosi+etal:underreview} under review). 
In a weighed generative LOT a representation's complexity is a function of its derivation cost.

@@

There are many ways to define priors over lexica (see, e.g., \citealt{goodman+etal:2008,
  piantadosi+etal:2012a}) but the key assumption here, common to all of them, is that simple
representational expressions should be favored over more complex ones. We should stress that
these details -- from the generative grammar to its complexity measure -- are to be regarded as
one convenient operationalization of one general approach to explicating learning biases; this
is not a commitment that this general approach is necessarily superior or that, within it, this
particular instrumentalization is the single most plausible.

@@

Recall that there are three parameters: soft-max parameter $\lambda$ affects how strongly
speakers favor messages that appear best from their subjective point of view; the
bottleneck size $k$ influences how faithfully learners can identify their teacher type; $l$
defines the learners' disposition towards choosing the most likely teacher type from the
posterior distribution. We expect that competitor types (types with lexica of the kind
$\Lbound$) have a fitness advantage over target types (pragmatic agents with lexica of the kind
$\Llack$), especially for very low levels of $\lambda$. Selection based on fitness alone may
therefore not lead to prevalence of target types in the population. On the other hand, lexica
of type $\Llack$ are simpler than those of type $\Lbound$ by the postulated measure from
above. This may make them more likely to be adopted by learners, especially when $k$ is low so
that different teacher types are relatively indistinguishable based on their behavior and when
$l$ is high. Still, lexica of the kind $\Lall$ are in turn even more likely \emph{a priori} than
lexica of the kind $\Llack$. Simulation results will shed light on the question whether target
types can emerge, and for which parameter constellations.

To better understand the workings of a pressure towards communicative efficiency and a pressure
towards learnability, we look at the behavior of the replicator and mutator step first in
isolation, and then in combination. All simulation runs are initialized with an arbitrary
distribution over types, constituting a population's first generation. All reported results are
the outcome of 50 update steps. These outcomes correspond to developmental plateaus in which
change is, if not absent, then at least very slow. In other words, even if the resulting states
do not correspond to an eventual attracting state, they characterize almost stationary states
in which the system remains for a very long time. As specified in \S\ref{sec:learnability}, the
mutation matrix $Q$ can be obtained by considering all possible state-message sequences of
length $k$. Given that this is intractable for large $k$, the sets of data which learners are
exposed to are approximated by sampling $250$ $k$-length sequences from each type's production
probabilities.

@@

\begin{figure}[t]
\centering
\includegraphics[width=\textwidth,height=8cm, keepaspectratio]{./plots/alt-fig1-onlyr}
\caption{Proportion of target types, pragmatic competitor types, pragmatic $\Lall$, and the population's majority type, in representative populations after $50$ generations under only a
  pressure for communicative success.}
\label{fig:only-R}
\end{figure}

Selection based on communicative success is sensitive to $\lambda$ since $\lambda$
influences signaling behavior, which in turn determines communicative success. This is
showcased in Fig. \ref{fig:only-R}, which shows the proportion of target types, pragmatic competitor types and pragmatic $\Lall$ in a representative population after $50$ replicator steps. The plot also indicates the proportion of the
\emph{majority type}: the type with the highest proportion in the final population. With
low $\lambda$ many types have very similar behavior, so that evolutionary selection lacks grip
and becomes very slow. The result is a very long transition with near stagnancy in a rather
homogeneous population with many types. Conversely, higher $\lambda$ promotes 
less-stochastic pragmatic behavior, widening the gap in expressivity between types and
promoting more homogeneous populations. As suggested by Fig. \ref{fig:only-R}, the majority
in most populations is not one of the six pragmatic $\Llack$-style types. That is, a pressure
only for communicative success does not lead to a prevalence of target types under any
$\lambda$-value. For instance, with $\lambda = 20$ $1000$ independent populations only had $11$
cases in which the target type was the majority type, corresponding to a mean proportion of
$.003$ across populations. By contrast, in $913$ cases the majority types had $\Lbound$ with
close to an even share between literal ($454$) and pragmatic types ($459$), corresponding to a
mean proportion of about $.48$ taken together.

@@

The effect of iterated learning without a pressure for communicative success using either posterior
sampling ($l = 1$) or a stronger tendency towards posterior maximization ($l = 15$) is shown in
Fig.\ref{fig:only-M} together with the prior over types. The prior shows that while users of
$\Llack$ are not the most favored by the inductive bias (compared, e.g., to $\Lall$) they are
nevertheless more advantaged than others, such as $\Lbound$, in virtue of the relatively simple
semantics they conventionalize (see Section~\ref{sec:an-induct-learn}). Crucially, $\Llack$
enables its users to convey each state with a single message when combined with pragmatic
reasoning and sufficiently high $\lambda$. This makes it less likely to be confused with
other types if the learning data is not too sparse ($k \geq 5$). Put differently, learners have
a higher propensity to infer pragmatic $\Llack$ when the teacher's type produces very similar
data, such as when using $\Lbound$. Moreover, $\Llack$ is less likely to be confused with
types with different observable behavior because its pragmatic use approximates a one-to-one
form-meaning mapping. As a consequence, a stronger propensity to maximize the posterior
increases their proportion in the population.

@@

As before, low $\lambda$ and $l$ lead to the prevalence of communicatively suboptimal types
that are a priori favored, such as $\Lall$. An increase in $\lambda$ leads to the selection of
target types but does not lead to monomorphic populations if learners sample from the
posterior (see the uppermost row in Figure \ref{fig:rmd}). Finally, a combination of high $\lambda$ and $l$ leads to increasing proportions of
a single majority target type. This joint influence is summarized in Fig. \ref{fig:diff}, which shows the mean difference between the highest target type and the highest proportion of a type of a different kind in $1000$ independent populations across $\lambda$ and $l$ values. Higher values of $\lambda$ and $l$ increase the prevalence of a single target type, whereas lower values lead to less pronounced differences, with a valley resulting from low $\lambda$ and high $l$ (cf. Fig. \ref{fig:rmd} with $\lambda = 1$ and $l=15$).

>>>>

\begin{example}
  Consider a simple and abstract coordination game. Agents are of two types: positive or negative. If
  agents of different type interact with each other, they obtain a payoff of $0$. If negative meets
  negative, each receives a payoff of $1$. If positive meets positive, they get a payoff of $2$.  A population state is completely characterized by the proportion $x$ of negatives. The fitness
  of negatives in population state $x$ is $f_n(x) = x$, that of positives is $f_p(x) = 2-2x$.
  The average fitness is $\Phi(x) = x f_n(x) + (1-x) f_p(x) = 3x^2 - 4x + 2$. Without mutation,
  the replicator dynamic will update $x$ to $\text{RD}(x) = \nicefrac{f(x)x}{\Phi(x)} = \nicefrac{x^2}{\Phi(x)}$. The update
  function $\text{RD}(x)$ of the replicator step is plotted in Fig.~\ref{fig:Updates_RMS} as the
  blue line. Rest points, for which $\text{RD}(x)=x$, are at $x=0$, $x=1$ and $x= \nicefrac{2}{3}$.
  The former are attractors as nearby points converge to them. Points near $x=\nicefrac{2}{3}$
  also move towards 0 or 1. This is schematically pictured in the topmost phase portrait in
  Fig.~\ref{fig:Phase_RD}. Adding mutation changes the dynamic and its rest points. Let us assume that $Q_{ji} = .9$
  when $j=i$. This is the proportion of types that are replicated faithfully. Conversely, a proportion of $.1$ will change their type from positive to negative, and vice-versa. The update effect of mutation on its own are described by $\text{M}(x) = .9 x + .1 (1-x) = .8x + .1$, plotted as the linear green line in Fig.~\ref{fig:Updates_RMS}. As shown in Fig.~\ref{fig:Phase_RD}, in this example mutation alone has only one stable rest point. It is located at $x = .5$. If we first take the replicator step and
  then the mutation step in sequence, we obtain the replicator mutator dynamic
  $\text{RMD}(x) = \text{M}(\text{RD}(x)) = \nicefrac{.9x^2 - .2x +.2}{3x^2 - 4x + 2}$, which is plotted in red in
  Fig.~\ref{fig:Updates_RMS}. The rest points are at $x=.121$, $x=.903$ and $x=.609$. The
  former two are attractors (see Fig.~\ref{fig:Phase_RD}).
\end{example}

@@

\subsection{Example}\label{subsec:example}
Section~\ref{sec:si-case-study} shows how these components and their interaction can shed light on the evolution of scalar implicatures. Before moving on, we will illustrate the model's workings in a constrained type space in order to showcase the effect that pressure for communicative success and learnability have on types. To this end, we focus only on the four types from Section~\ref{sec:languages+use}. These types are made up of a combination of either lexicon $\Llack$ or $\Lbound$ and either literal (level-$0$) or pragmatic (level-$1$) behavior.

@@

A population will be made up of frequencies of these four types. We may think about such a population as a point in a two-dimensional space. In Fig.~\ref{fig:sketch} this space is outlined by the grey dotted square. The first dimension determines the proportion of types that lexicalize ambiguous $\msome$ (the $x$-axis in Fig.~\ref{fig:sketch}). That is, the proportion of types with $\Llack$. The second dimension determines the proportion of pragmatic language users (the $y$-axis in Fig.~\ref{fig:sketch}). Populations at the either corner of this space are monomorphic, consisting purely of a single type. If we move from away from edges toward the center of the space, then this two-dimensional representation underspecifies the proportion of types found in a population. The space's center, for instance, could be composed 50\% of pragmatic $\Llack$ and of an equal proportion of literal $\Lbound$; but it could equally well be a population that has each type with the same frequency. This will be sufficient for our purposes, as we are only interested in showcasing the rough trajectories of populations through this space after replicator or mutator steps. 
%The space's corners (and edges) are unequivocal: the lower-left corner corresponds to an population of only literal $\Lbound$ users; the lower-right corner to all pragmatic $\Lbound$ users; the upper-left to literal $\Llack$ users only; and the upper-right corner to the majority all pragmatic $\Llack$ users.

From Section~\ref{sec:languages+use}, we already know some things about this type space. When it comes to communicative success, types with $\Lbound$ have a functional advantage over $\Llack$; literal $\Lbound$ has a slight advantage over pragmatic $\Lbound$; all have a sizable advantage over literal $\Llack$. The red arrows in Fig.~\ref{fig:sketch} sketch out these relationships. They depict the rough trajectories we may expect if there is only pressure for communicative success. How much each replicator step moves a population ultimately depends on the rationality parameter $\lambda$. In particular, functional differences decrease as $\lambda$ increases: if $\lambda$ is sufficiently high, all types except for literal $\Llack$ speakers will show a strong tendency to associate $\ssome$ with $\msome$ and $\sall$ with $\mall$. As a consequence, when it comes to learnability and transmission fidelity, the likelihood that any of the three remaining types produces a given datum $d$ will be approximately equal. In other words, the likelihood offers little differentiation for learners if $\lambda$ is high. We can therefore expect that, if the prior is flat, no type is passed on very faithfully. This would lead populations to gravitate toward the space's center after mutator steps. This is sketched by the blue dash-dotted arrows in Fig.~\ref{fig:sketch}. Neither prior nor posterior alone, but their combination, are what determines the effects of (iterated) Bayesian learning. If, instead of a flat prior, there is a prior that favors, say, $\Llack$ over $\Lbound$ we can expect mutator steps to be skewed toward the upper-half of the space. This is suggested by the purple arrows in Fig.~\ref{fig:sketch}. Taking stock, the effects of iterated learning will depend not only on $\lambda$, affecting the likelihood, but also on the prior, how strong the bias is, and the degree to which learners tend to maximize the posterior, controlled by parameter $l$.

Fig.~\ref{fig:quiver} shows actual trajectories, showcasing how they are modulated by parameter values. The replicator steps make the functional disadvantage of $\Llack$ against $\Lbound$ apparent, particularly for literal types and/or low $\lambda$. They also highlight that the functional difference between literal and pragmatic $\Lbound$ is small. Under functional pressure only, populations move toward the lower part of the square but the steps become small as the edge, representing populations of only $\Llack$, are reached. As for the mutator diagrams, there are two main things to note. As expected, a small bias for a type will lead to larger proportions to adopt it provided that likelihood offers little to no differentiation between types. This is the situation the right-most space in Fig.~\ref{fig:quiver} depicts. However, the bias is far from absolute. If, instead, certain types are more likely to be recovered based on the linguistic input they produce, they will be passed on more faithfully. This is why, for instance, populations move toward a majority of $\Llack$-types for $\lambda = 5$ and $l = 10$.

@@

Fig.~\ref{fig:sketch} and \ref{fig:quiver} depict replicator and mutator steps in isolation to better showcase their effects. Their combination will expectedly yield different outcomes, which we now turn to in a case study on scalar implicatures.

@@

The model of the previous section formalizes the evolutionary competition between different
sets of lexical representations and ways of using them. This section looks at a case study on
scalar implicatures. It engages in a formal thought experiment to address the
question: if a population of language users could freely combine different lexica with
different dispositions to act on them, what are conditions under which the majority view of scalar
implicatures could have evolved?

@@

Consider a state space with three states $\States = \set{\snone, \ssome, \sall}$ and think of
it as a partition of possible worlds into cells where none, some or all of the $A$s are $B$s,
for some arbitrary fixed predicates $A$ and $B$. Enlarging the setup from Section~\ref{sec:languages+use} by another state will enable us to discern the difference between three kinds of lexica we will use to unravel the model's dynamics (see below). Eight lexical representations can be distinguished based on their truth or falsity in three world states, six of which are not contradictory or tautological (see Table~\ref{tab:concepts} below).

@@

Recall that types are a combination of a lexicon and a manner of language use. We analyze the
model's predictions in populations of types with one of the two behaviors introduced earlier:
literal or pragmatic. The former correspond to level-$0$ reasoners and the latter to level $1$. Accordingly, there is a total of $432$ types. Six types are variants of pragmatic language
users with $\Llack$-like lexica. We refer to these as \emph{target types} because they
represent lexica and language use that conform to the majority view of scalar
implicatures. Twelve types are either literal or pragmatic users of lexica of the
$\Lbound$ kind. We refer to these as \emph{competitor types}, because they are expected to be
the target types' main contenders in evolutionary competition. Finally, note that while
different types may lexicalize the same representations, they may nevertheless map different
states to different overt messages. More informally, they speak different languages that
lexicalize the same concepts. Consequently, more often than not different types of the same
kind fail to understand each other.

@@

There is a growing effort to develop empirically testable representational languages that allow
for the measure of semantic complexity. For instance, so-called {\em languages of thought}
(LOTs) have been put to test in various rational probabilistic models that show encouraging
results (see, e.g., \citealt{katz+etal:2008}, \citeauthor{piantadosi+etal:underreview} under review; 2012, and \citealt{piantadosi+jacobs:2016} for recent discussion). At its
core, a LOT defines a set of operations and composition rules from which lexical representations can be
derived. As a first approximation and for the sake of concreteness, we follow this approach to
motivate and formalize a preference of learners for simpler semantic representations
(\citealt{feldman:2000, chater+vitanyi:2003, piantadosi+etal:2012a,kirby+etal:2015}, \citeauthor{piantadosi+etal:underreview} under review). 
In a weighted generative LOT a representation's complexity is a function of its derivation cost.

@@

The menu of inductive biases argued to shape language acquisition is steadily being refined. Apart from simplicity, prominent examples include mutual exclusivity \citep{merriman+bowman:1989,clark:2009}, regularization \citep{hudson+etal:2005}, and generalization \citep{smith:2011}. Even viewed in isolation, there are still many ways to define priors over lexica (see, e.g., \citealt{goodman+etal:2008, piantadosi+etal:2012a, kirby+etal:2015}). The key assumption here, common to all ways to define simplicity-biased priors, is that simple representational expressions should be favored over more complex ones. However, in view of the menu of possible operationalizations, other biases and their potential interplay, we should stress that these details -- from the generative grammar to its complexity measure -- are to be regarded as one convenient operationalization of one general approach to explicating learning biases; this is not a commitment that this general approach is necessarily superior or that, within it, this particular instrumentalization is the single most plausible. We return to this issue in relation to our results in Section~\ref{sec:an-induct-learn}.

@@

Recall that there are three parameters, which we hold constant across types: soft-max parameter $\lambda$ affects how strongly
speakers favor messages that appear best from their subjective point of view; the
bottleneck size $k$ influences how faithfully learners can identify their teacher type; $l$
defines the learners' disposition towards choosing the most likely teacher type from the
posterior distribution. We expect that competitor types (types with lexica of the kind
$\Lbound$) have a fitness advantage over target types (pragmatic types with lexica of the kind
$\Llack$), especially for very low levels of $\lambda$ (see Fig~\ref{fig:sketch}). Selection based on fitness alone may
therefore not lead to prevalence of target types in the population. On the other hand, lexica
of type $\Llack$ are simpler than those of type $\Lbound$ by the postulated measure from
above. This may make them more likely to be adopted by learners, especially when $l$ is high. Still, lexica of the kind $\Lall$ are in turn even more likely \emph{a priori} than lexica of the kind $\Llack$. As showcased in Section~\ref{subsec:example}, the prior is not all that counts for learnability. Simulation results will shed light on the question whether target
types can emerge, and for which parameter constellations.

To better understand the workings of a pressure towards communicative efficiency and a pressure
towards learnability, we look at the behavior of the replicator and mutator step first in
isolation, and then in combination. All simulation runs are initialized with an arbitrary
distribution over types, constituting a population's first generation. All reported results are
the outcome of 50 update steps. These outcomes correspond to developmental plateaus in which
change is, if not absent, then at least very slow. In other words, even if the resulting states
do not correspond to an eventual attracting state, they characterize population states in which the system remains for a long time. As specified in Section~\ref{sec:learnability}, the
mutation matrix $Q$ can be obtained by considering all possible state-message sequences of
length $k$. Given that this is intractable for large $k$, the sets of data which learners are
exposed to are approximated by sampling $250$ $k$-length sequences from each type's production
probabilities.

@@

\begin{figure}[t]
\centering
\includegraphics[width=\textwidth,height=8cm, keepaspectratio]{./plots/alt-fig1-onlyr}
\caption{Proportion of target types, \hl{pragmatic} competitor types, \hl{pragmatic} $\Lall$, and the population's majority type, in representative populations after $50$ generations under only a
  pressure for communicative success. \hl{Flip plots, change plot generation to show the highest 6 Lbound users and the highest Lall, irrespective of ling. behavior, change text surrounding these things}}
\label{fig:only-R}
\end{figure}

Selection based on communicative success is sensitive to $\lambda$ since $\lambda$
influences signaling behavior, which in turn determines communicative success. This is
showcased in Fig. \ref{fig:only-R}, which shows the proportion of target types, \hl{pragmatic} competitor types and \hl{pragmatic} $\Lall$ in a representative population after $50$ replicator steps. The plot also indicates the proportion of the
\emph{majority type}: the type with the highest proportion in the final population. With
low $\lambda$ many types have very similar behavior, so that evolutionary selection lacks grip
and becomes very slow. The result is a very long transition with near stagnancy in a rather
homogeneous population with many types. Conversely, higher $\lambda$ promotes 
less-stochastic linguistic behavior, widening the gap in expressivity between some types and
promoting more homogeneous populations. As suggested by Fig. \ref{fig:only-R}, the majority
in most populations is not one of the six pragmatic $\Llack$-style types. That is, a pressure
only for communicative success does not lead to a prevalence of target types under any
$\lambda$-value. For instance, with $\lambda = 20$ $1000$ independent populations only had $11$
cases in which the target type was the majority type, corresponding to a mean proportion of
$.003$ across populations. By contrast, in $913$ cases the majority types had $\Lbound$ with
close to an even share between literal ($454$) and pragmatic types ($459$), corresponding to a
mean proportion of about $.48$ taken together.

@@

The effect of iterated learning without a pressure for communicative success using either posterior
sampling ($l = 1$) or a stronger tendency towards posterior maximization ($l = 15$) is shown in
Fig.~\ref{fig:only-M} together with the prior over types. The prior shows that while users of
$\Llack$ are not the most favored by the inductive bias (compared, e.g., to $\Lall$) they are
nevertheless more advantaged than others, such as $\Lbound$, in virtue of the relatively simple
semantics they conventionalize (see Section~\ref{sec:an-induct-learn}). Crucially, $\Llack$
enables its users to convey each state with a single message when combined with pragmatic
reasoning and sufficiently high $\lambda$. This makes it less likely to be confused with
other types if the learning data is not too sparse ($k \geq 5$). Put differently, learners have
a higher propensity to infer pragmatic $\Llack$ when the teacher's type produces very similar
data, such as when using $\Lbound$. Moreover, $\Llack$ is less likely to be confused with
types with different observable behavior because its pragmatic use approximates a one-to-one
form-meaning mapping. As a consequence, a stronger propensity to maximize the posterior
increases their proportion in the population.

@@

As before, low $\lambda$ and $l$ lead to the prevalence of communicatively suboptimal types
that are a priori favored, such as $\Lall$. An increase in $\lambda$ leads to the selection of
target types but does not lead to monomorphic populations if learners sample from the
posterior (see the uppermost row in Fig.~\ref{fig:rmd}). Finally, a combination of high $\lambda$ and $l$ leads to increasing proportions of
a single majority target type. This joint influence is summarized in Fig. \ref{fig:diff}, which shows the mean difference between the highest target type and the highest proportion of a type of a different kind in $1000$ independent populations across $\lambda$ and $l$ values. Higher values of $\lambda$ and $l$ increase the prevalence of a single target type, whereas lower values lead to less pronounced differences, with a valley resulting from low $\lambda$ and high $l$ (cf. Fig. \ref{fig:rmd} with $\lambda = 1$ and $l=15$).
